---
editor_options: 
  markdown: 
    wrap: sentence
---

### Exercise 0 : System prompt

You are helping me with an advanced statistical programming course in R.

Code style: - Use the tidyverse (dplyr, tidyr, ggplot2, purrr) where possible.
- Use the native pipe \|\> rather than %\>% unless it requires more complicated code.
- Follow the tidyverse style guide: snake_case, \<- for assignment, two-space indent.
- Prefer small, pure functions over long scripts.
- For file paths use the `here` package.

When I ask for help debugging: - First explain what the error message means, in plain English.
- Then suggest where to look in my code, BEFORE proposing a fix.
- Only write a full rewrite if I ask for one.

General: - Do not apologise.
Do not pad with filler.
If you are uncertain, say so.
- If I paste in code, assume it is real code I care about — do not silently rename variables or change the pipe style.

### Exercise 1

```{r}
library(tidyverse)
```

1.  buggy code:

    ```{r}
    library(palmerpenguins)
    library(dplyr)

    mean_mass_by <- function(data, group_var) {
      data |>
        group_by(group_var) |>
        summarise(mean_mass = mean(body_mass_g, na.rm = TRUE))
    }

    mean_mass_by(penguins, species)
    ```

2.  The intention of the function is probably to give out a vector of the mean mass of a number of subgroups created by a categorical variable.

    Possible bugs may be that the group_by() doesn t count since there is no return given and therefore only the last calculation might be done

    or that there have to be "" somewhere

3.  full error message :

    Error in \`group_by()\`: !
    Must group by variables found in \`.data\`.
    ✖ Column \`group_var\` is not found.
    Backtrace:

    1.  global mean_mass_by(penguins, species)
    2.  dplyr:::group_by.data.frame(data, group_var)

4.  Asking LLM about error message, template :

    Here is my code: `[code]`.
    Here is the error: `[error]`.
    Please explain what R is complaining about, and point me to the line where the problem originates.
    Do not rewrite my code yet.

    Answer:

    R is complaining that `group_by()` is looking for a column literally named `group_var`, and it cannot find one in `data`.

    In plain English:

    -   Inside your function, `group_var` is a function argument.

    -   But `dplyr::group_by()` is interpreting `group_var` as a column name in the data frame.

    -   Your `penguins` data has a column called `species`, not a column called `group_var`.

    -   So `group_by()` throws:

    > Must group by variables found in `.data`\
    > Column `group_var` is not found.

    Where the problem originates:

    -   The key line is inside your function:

    ```         
    data |> group_by(group_var) 
    ```

    That is the line causing the error.

    Why the backtrace points there:

    -   The call `mean_mass_by(penguins, species)` goes into your function.

    -   Then `group_by(group_var)` runs.

    -   `group_by()` searches the data for a column named `group_var`.

    -   It does not find one, so it errors.

    One more thing to notice:

    -   In the code you first pasted, you wrote `mean_mass_by(penguins, gruspecies)`, which would also be a problem because `gruspecies` is not a valid column name.

    -   But the actual error/backtrace shows the call was `mean_mass_by(penguins, species)`.

    -   So the real issue is not the column `species`; it is that the function is not passing that argument into `group_by()` in the way `dplyr` expects.

    So: start by looking at the `group_by(group_var)` line inside `mean_mass_by()`.
    That is where the mismatch begins.

5.  Fix the code yourself you lazy dog!

    ```{r}

    #curly brackets {{}} around the variable inside the calculation makes it a variable in r and not just a string


    mean_mass_by <- function(data, group_var) {
      
     
      data |>
        group_by({{group_var}}) |>
        summarise(mean_mass = mean(body_mass_g, na.rm = TRUE))
    }

    mean_mass_by(penguins, species)


    ```

### Exercise 2: Locating errors

1.  buggy code:

    ```{r}
    #library(palmerpenguins)
    #library(dplyr)
     
    summarise_species <- function(data) {
      data |>
        group_by(species) |>
        summarise(mean_mass = mean_body_mass(body_mass_g))
    }
     
    mean_body_mass <- function(x) {
      mean(x, na.rm = TRUE)   # error was here - typo TREU instead of TRUE
    }
     
    summarise_species(penguins)
    ```

2.  ggg

3.  locate error with rlang::last_trace()

    ```{r}
    rlang::last_trace()
    ```

4.  put another wrong value

```{r}
summarise_species <- function(data) {
  data |>
    group_by(species) |>
    summarise(mean_mass = mean_body_mass(body_mass_g))
}
 
mean_body_mass <- function(x) {
  mean(x, na.rm = banana)   
}
 
summarise_species(penguins)
```

5.  fix it with TRUE

### Exercise 3: browser()

1\.

```{r}
my_factorial <- function(n) {
  
  stopifnot (n >= 0, n == as.integer(n))
  if (n <= 1) return (1)
  return (n*my_factorial(n - 1))
}

my_factorial(0)   


```

```{r}

my_factorial <- function(n) {
  stopifnot (n >= 0, n == as.integer(n)) 
  if (n <= 1) return(1)
  return(n * my_factorial(n - 1))
}

my_factorial(0)
```

### exercise 4: debug () , debugonce()

```{r}
standardise <- function(x) {
  (x - mean(x, na.rm = T)) / sd(x, na.rm = T)
}

standardise(c(1, 2, 3, 4, 5))   # fine
standardise




)  # returns all NAs — why?





```

1\.

2\.
mean and sd return NA if one value is NA

3\.
I would use debugonce() if I already have a good idea what the problem is and debug if I want to keep looking at the function longer

### ex 5: debugging quarto
